Learning with Mental Models: the case of overestimation

### 2023-10-25 Paper/draft.Rmd
added last session

---
title: "Learning with Mental Models: the case of overestimation"  
date: '`r format(Sys.time(), "%B %d, %Y")`'
author: "Jimena Galindo"
abstract: "I design a framework and a laboratory experiment that allows for the comparison of multiple theories of misspecified learning. I focus on a framework with endogenous information and a data-generating process ruled by two fundamentals: an ego-relevant parameter and a state. Within this framework I study three forces that can lead to misspecified beliefs: initial misspecifications, learning traps and biased updating. I find that biased updating is the main driver of misspecified beliefs in the lab. In addition, I vary the degree of ego-relevance of the parameter by introducing a stereotype treatment. The data is consistent with biased updating in both cases but for potentially different reasons: when learning about themselves, subjects attribute successes to their own ability and failures to luck. Instead, in the stereotype treatment, they compensate for initial negative biases by over-attributing positive signals to the ability of others. This translates into similar observed choices but different dynamics in beliefs."

@@

In all of these examples,
holding an incorrect belief about a fundamental leads to suboptimal choices with 
potentially high costs. In spite of the abundance of evidence, the scope of the existing research 
is limited in terms of the frameworks it considers. Most of the experimental evidence 
documenting the bias is collected in settings where subject are passive learners.\footnote{
\citet{Gotte2022} and \citet{Ozyilmaz2022} are exceptions that study settings with endogenous information processes.} They 
observe a noisy signal and report their beliefs over subsequent rounds.\footnote{\citet{Bracha2012} and 
\citet{Mobius2022} are some examples.} Although these studies provide important insights, they are 
not flexible enough 
to incorporate the richer theories that have been proposed more recently. In particular, they 
do not allow the study of endogenous learning or for learning about multiple parameters at once.\footnote{\citet{Coutts2020} 
studies an environment with an ego-relevant parameter and an exogenous state but does not 
incorporate the endogenous information process.}

@@

When the setting features learning traps, even an agent who incorporates all information correctly, 
may fall into learning traps as outlined by @Hestermann2021. These traps are characterized by a 
combination of an incorrect belief and an
optimal action which produce information that confirms the incorrect belief. 
Once an agent falls into a trap, the belief will be stable and even with 
a correctly specified model of learning, they will not be able to abandon their misspecified beliefs. 
If, the agent is 
dogmatic about their initial belief, @Heidhues2018 show that they will inevitably fall into a trap
and thus will be able to rationalize and sustain their initial misspecification belief.\footnote{\citet{Gotte2022} study the case of agents with dogmatic initial beliefs in a 
laboratory experiment.}

@Ba2023 moves away from dogmatism and endows the agents with a mechanism through which 
they can abandon incorrect beliefs; this allows them to avoid falling into learning traps. 
To do so agents perform Bayesian hypothesis tests that evaluate which is the more 
likely parameter out of two possibilities. By doing this, she characterizes the set of situations 
in which, even agents who consider alternative paradigms, may become trapped.\footnote{A 
similar mechanism is proposed by \citet{Schwarstein2021} in a setting with persuasion.}

@@

The theories that I consider make predictions in two distinct frameworks. I first detail each theory 
within their original framework and then develop a unifying example that allows me to
compare the predictions of all of them. In framework 1 I consider a setting in which the an agent observes
a continuous output and infers the underlying state. In framework 2, the agent observes a binary 
outcome and infers the underlying probability of success. In both frameworks, I consider two nested theories 
of learning.

@@

The agent chooses an action $a\in A$ and observes a noisy outcome $h$. 
The outcome is a function of the agent's type, the state, and
the action. In particular $h = h(\theta^*, \omega^*, a) + \varepsilon$ with $h(\cdot)$ increasing in both $\theta^*$ and $\omega^*$, 
and such that conditional on a pair of parameters $(\theta, \omega)$, there is a unique optimal action. 
additionally, $\varepsilon\sim N(0, \sigma)$ is noise in the output.

After observing the outcome, the agent updates his beliefs about $\theta$ and $\omega$ using some algorithm and moves on to the next period.
He repeats this process infinitely many times. I make the simplifying assumption that the agent is myopic and chooses the action 
that maximizes the payoff in each period. This assumption simplifies the analysis and plays a role in whether an agent who updates
 their beliefs using Bayes rule would learn the truth or not, however, for the main theory discussed in this section, the results hold
 even when relaxing this assumption. The behavior of a forward-looking agent is further discussed in the conclusion.

A key notion in this setting is that of a self-defeating equilibrium\footnote{This notion is an adaptation of the Berk-Nash equilibrium in 
\citet{Esponda2016} to this setting with only one agent}. 
A \emph{self-defeating equilibrium} is a belief and action pair such that the agent's belief about their type is misspecified and the 
outcome generated by the action is consistent with the misspecified belief. This means that the average outcome under the true type and the true 
state equals the average output the agent expects under the misspecified belief. The agent's belief is said to be 
\emph{stable} when this happens.

Within this framework, I consider two nested theories of belief updating: the first one is a dogmatic modeler from @Heidhues2018,
The second one is a switcher, as in @Ba2023. The dogmatic modeler can be seen as a switcher with an infinitely sticky initial belief
so the two theories are nested. However, outside of the limiting case, they produce different 
predictions about the agent's behavior. I discuss each of them separately in what follows.

@@

Take the action space to be the same as the state space $\Omega$ and  $\Omega =A=[\underline{a}, \infty)$.

A dogmatic agent does not update their beliefs about $\theta$, instead, he holds a degenerate belief that places probability on 
$\hat{\theta}$, which is potentially misspecified. 
In this case, no matter
how much information he gathers against being of type $\hat{\theta}$, he will not update his beliefs about it. 
Any discrepancies between the 
observed outcomes and his believed type are incorporated using the Bayes rule to update his beliefs about $\omega$.
@Heidhues2018 show that, under certain assumptions on the per-period utility,\footnote{The assumptions are that 
$u$ is twice continuously differentiable with: (i)$u_{aa}<0$ and $u_a(\underline{a},  \theta, \omega)>0>u_a(\Bar{a}, \theta, \omega)$, 
(ii) $u_{\theta}, u_{\omega}>0$ and (iii) $u_{a\theta}<0$ and $u_{a\omega}>0$. Where $\underline{a}$ is the
minimal action and $\Bar{a}$ is the maximal one. 
 The direction of the derivatives is a normalization
and the results would hold even when the signs are reversed.} 
a dogmatic modeler will inevitably fall into a self-defeating equilibrium. The equilibrium will be such that
the outcomes they observe reinforce their belief on $\omega$ in such a way that as $t\to\infty$ the agent will 
be sure that
the state is some $\omega^{'}$ consistent with their believed type and the observed data. In other words, 
they will be in a self-defeating equilibrium with a stable belief that places probability one on the incorrect
parameters $(\hat{\theta}, \omega_{\infty})$

The mechanism by which the dogmatic agent falls into the self-defeating equilibrium is the following: Suppose the agent holds the 
misspecified belief that they are type $\hat{\theta}>\theta^*$. For any prior over $\omega$, the agent will be disappointed by the outcome.
He expected a gain of $h(\hat{\theta}, \mathbb{E}(\omega), a)$ but instead observes  $h(\hat{\theta}, \omega^*, a)$. 
There are two 
possible sources for the disappointment: the first is that the realized state is lower than the expected state; the second source is that the
agent is of type $\theta^*$ and therefore, for all possible states, his gain will be lower than what he expected. Because the agent is dogmatic,
he will not update his beliefs about $\theta$ and as a consequence will attribute the disappointment to the state being lower than expected.
He will continue to update in this way until he converges to a belief about $\omega$ that is stable. Such a belief will 
explain the observed utility perfectly and allow the agent to rationalize his dogmatic belief about $\theta$. 
Under the assumptions of @Heidhues2018, there is a unique value of $\omega$ at which the belief is stable, I will refer to such value
as $\omega_\infty$.
This mechanism is illustrated in Example 1.

@@

Where $h^t$ is the history of outcomes up to time $t$ and $\alpha$ is the switching threshold.\footnote{Notice that if $\alpha \to \infty$, the behavior of the switcher will be indistinguishable from that of the Dogmatic modeler. 
In this sense, the switcher is a generalization of the dogmatic type.} By keeping track of the 
likelihood ratio, the agent can perform a \emph{Bayesian hypothesis test} and adopt the Dogmatic belief that best fits 
the data.\footnote{In a related problem \citet{Schwarstein2021} proposes a similar updating procedure which relies on the Bayesian 
hypothesis test. However, in their model there is a sender who optimally chooses to propose a model that fits the data}

@@

A Bayesian agent simultaneously updates their beliefs about $\theta$ and $\omega$ by using Bayes' rule. 
The posterior at period t+1 about $\theta$ after observing outcom $o$ is given by:
$$
p_{t+1}(\theta, \omega| h^t) = \frac{p[o|\theta, \omega]p_t(\theta, \omega)}{\sum_{(\theta', \omega')}p[o|\theta', \omega']p_t(\theta', \omega')}
$$

@@

Bayesian agents choose the effort level that maximizes their expected flow payoff by taking expectations over
their prior beliefs about $\theta$ and $\omega$. Since agents are myopic, even though all the parameters 
could be identified with enough variation in choices, there will be instances in which the Bayesian will not
learn even with infinite amounts of data. This happens because, by being myopic, they do not internalize 
the tradeoff between flow payoff
and learning. This can result in too little experimentation to learn their true type. An alternative to this approach is given 
by @Hestermann2021 and is discussed in the concluding remarks. Regardless, this approach is useful as a 
benchamrk for the other theories discussed in this paper.

@@

A Self-Serving Bayesian is an agent who uses a biased version of Bayes rule to update their beliefs. 
They will update their beliefs 
about the state $\omega$ and his type $\theta$ simultaneously
by over-attributing successes to a high value of $\theta$ and under-estimating the role of higher $\omega$. Similarly,
he will attribute failure to a low state to a greater degree than an unbiased agent would. 
To model the self-serving attribution bias, I take the approach of @benjamin2019, where the posterior odds are given by:

@@

with $c(\theta_H, \omega, o)<c(\theta_M, \omega, o)<c(\theta_L, \omega, o)\leq 1$
and $c(\theta, \omega_L, o)<c(\theta, \omega_M, o)<c(\theta, \omega_H, o)\leq 1$.

@@

@@

In order to compare the predictions of the theories discussed above, I develop a unifying example that allows me to
isolate the forces behind each of the theories. The example is a modification of the one in @Heidhues2018 and is adapted to 
implementable in the lab.

@@

Denote the true type by $\theta^*$ and the true state by $\omega^*$. The agent holds some prior belief about 
$\theta$\footnote{which is potentially misspecified as in the dogmatic and switcher cases discussed above} 
and chooses a binary gamble $e in \{e_L, e_M, e_H\}$. The agent observes whether the gamble is a success or a failure and 
gets a payoff of $1$ and if it is a success; they get $0$ otherwise.

@@

In this section, I outline the behavior that is predicted by each of the theories discussed above.

@@

@@

**Prediction 1S:** With positive probability, the switcher will escape the self-defeating equilibrium and learn the true state.

One caveat is that when the switcher and the dogmatic agent both start with a correctly specified belief, neither of them 
will fall into the self-defeating equilibria and thus will look identical even in the long run. This means that
in order to distinguish between the two theories, I need to look at cases where the agent starts with a misspecified belief.

@@

**Prediction 1A:** _After observing a failure, the agent will adjust their effort downwards by more than a Bayesian would have done.
After observing a success, the agent will adjust their effort upwards by less than a Bayesian would have done. If the bias is large enough,
the agent might not adjust their effort upwards at all, or even decrease their effort in response to a success._

Prediction 1A helps distinguish between the self-attribution bias and the dogmatic modeler. However, it does not help distinguish between
the self-attribution bias and the switcher. In order to do so, I need to look at cases where the agent starts with a correctly specified belief.
In these cases, the switcher is very unlikely to switch to a different belief and therefore, the behavior of the switcher will mostly change
in the same direction as the information he receives. \footnote{If they get a streak of failures they are likely to 
adjust their effort downwards and if they get a streak of successes they are likely to adjust their effort upwards.} Instead, 
the self-attribution bias will lead to a different behavior. In particular, even for agents who have a belief 
that is correctly specified and place a lot of weight on the correct values of $\theta$, the bias will 
cause these agents to become overconfident after a sequence of successes. This force is not present in the switcher or the dogmatic
and it is the key feature that allows me to identify a biased updating procedure.

@@

## The Bayesian

The Bayesian serves as the closest benchmark to a fully rational agent. However, as discussed above, 
they are still vulnerable to falling into self-defeating equilibria, in which case, their long-run behavior
will be identical to that of the dogmatic modeler. The main distinguishing feature is that when they do 
fall into a self-defeating equilibrium, there is no bias in the direction of the misspecification that they 
end up being trapped in. In contrast, the dogmatic modeler will always be trapped in a misspecified belief
that they started with and the self-attribution updater will always become overconfident.

**Prediction 1B:** _When an agent starts with a diffused prior and falls into a self-defeating equilibrium,
they will be equally likely to fall into the underconfident trap as the overconfident trap._

**Prediction 2B:** _Bayesian agents always update their effort choices in the same direction as the information they receive. 
That is, if they see successes, they would only increase their effort; and if they see failures they would 
only decrease their effort._

@@

$$ arg\min_j \sum_k \frac{1}{T}\sum_t{\left[e_{ikt}-\tilde{e}_{t}(\theta, \omega, j)\right]} $$
The histogram in Figure \ref{fig:best-models} shows the distribution of the best fitting models in the sample. The average fit of each model is reported in the Appendix.

@@

Another bias that has been studied and suggested in the literature is the confirmation bias. In particular,
it has been suggested that subjects who are overestimate $\theta$ in their prior, will over-attribute 
successes to a high type and under-attribute them to a low type. Symmetrically, subjects who initially 
underestimate $\theta$ will over-attribute failures to a low type and under-attribute them to a high type.
I do not find evidence of such a bias in the data. This is consistent with what @Mobius2022 find in their
experiment.

I further investigate possible heterogeneity in the underlying models being used by estimating the 
structural parameters of the models.

>>>>

---
title: "Learning with Mental Models: the case of overestimation"  
date: '`r format(Sys.time(), "%B %d, %Y")`'
author: "Jimena Galindo"
abstract: "I design a framework and a laboratory experiment that allow for the comparison of multiple theories of misspecified learning. I focus on a framework with endogenous information and a data-generating process ruled by two fundamentals: an ego-relevant parameter and a state. Within this framework I study three forces that can lead to misspecified beliefs: initial misspecifications, learning traps and biased updating. I find that biased updating is the main driver of misspecified beliefs in the lab. In addition, I vary the degree of ego-relevance of the parameter by introducing a stereotype treatment. The data is consistent with biased updating in both cases but for potentially different reasons: when learning about themselves, subjects attribute successes to their own ability and failures to luck. Instead, in the stereotype treatment, they compensate for initial negative biases by over-attributing positive signals to the ability of others. This translates into similar observed choices but different dynamics in beliefs."

@@

In all of these examples,
holding an incorrect belief about a fundamental leads to suboptimal choices with 
potentially high costs. In spite of the abundance of evidence, the scope of the existing research 
is limited in terms of the frameworks it considers. Most of the experimental evidence 
documenting the bias is collected in settings where subjects are passive learners.\footnote{
\citet{Gotte2022} and \citet{Ozyilmaz2022} are exceptions that study settings with endogenous information processes.} They 
observe a noisy signal and report their beliefs over subsequent rounds.\footnote{\citet{Bracha2012} and 
\citet{Mobius2022} are some examples.} Although these studies provide important insights, they are 
not flexible enough 
to incorporate the richer theories that have been proposed more recently. In particular, they 
do not allow the study of endogenous learning or for learning about multiple parameters at once.\footnote{\citet{Coutts2020} 
studies an environment with an ego-relevant parameter and an exogenous state but does not 
incorporate the endogenous information process.}

@@

When the setting features learning traps, even an agent who incorporates all information correctly, 
may fall into learning traps as outlined by @Hestermann2021. These traps are characterized by a 
combination of an incorrect belief and an
optimal action which produce information that confirms the incorrect belief. 
Once an agent falls into a trap, the belief will be stable and even with 
a correctly specified model of learning, they will not be able to abandon their misspecified beliefs. 
If the agent is 
dogmatic about their initial belief, @Heidhues2018 show that they will inevitably fall into a trap
and thus will be able to rationalize and sustain their initial misspecification belief.\footnote{\citet{Gotte2022} study the case of agents with dogmatic initial beliefs in a 
laboratory experiment.}

@Ba2023 moves away from dogmatism and endows the agents with a mechanism through which 
they can abandon incorrect beliefs. This allows them to avoid falling into learning traps. 
To do so agents perform Bayesian hypothesis tests that evaluate which is the more 
likely parameter out of two possibilities. By doing this, she characterizes the set of situations 
in which, even agents who consider alternative paradigms, may become trapped.\footnote{A 
similar mechanism is proposed by \citet{Schwarstein2021} in a setting with persuasion.}

@@

The theories that I consider make predictions in two distinct frameworks. I first detail each theory 
within their original framework and then develop a unifying example that allows me to
compare the predictions of all of them. In framework 1 I consider a setting in which the an agent observes
a continuous output and infers the underlying parameters. In framework 2, the agent observes a binary 
outcome and infers the underlying probability of success, which is related to the parameters. In both frameworks, I consider two nested theories of learning.

@@

The agent chooses an action $a\in A$ and observes a noisy outcome $h$. 
The outcome is a function of the agent's type, the state, and
the action. In particular $h = h(\theta^*, \omega^*, a) + \varepsilon$ with $h(\cdot)$ increasing in both $\theta^*$ and $\omega^*$, 
and such that conditional on a pair of parameters $(\theta, \omega)$, there is a unique optimal action. 
Additionally, $\varepsilon\sim N(0, \sigma)$ is noise in the output.

After observing the outcome, the agent updates his beliefs about $\theta$ and $\omega$ using some algorithm and moves on to the next period.
He repeats this process infinitely many times. I make the simplifying assumption that the agent is myopic and chooses the action 
that maximizes the payoff in each period. This assumption simplifies the analysis and plays a role in whether an agent who updates their beliefs using Bayes rule would learn the truth or not, however, for the main theory discussed in this section, the results hold even when relaxing this assumption.

A key notion in this setting is that of a self-defeating equilibrium\footnote{This notion is an adaptation of the Berk-Nash equilibrium in \citet{Esponda2016} to this setting with only one agent}. 
A \emph{self-defeating equilibrium} is a belief and action pair, $(p, a)$, such that the agent's belief about their type is misspecified, and the outcome generated by the action is consistent with the misspecified belief. The agent's belief is said to be \emph{stable} when this happens and the average output is

$$\bar{h}(a^*) = h(\hat{\theta}, \hat{\omega}, a)$$
Where $\hat{\theta}$ and $\hat{\omega}$ is the type-state pair on which the misspecified belief places all probability. That is, the agents belief is $p(\hat{\theta}, \hat{\omega}) = 1$ and $p(\theta, \omega) = 0$ for all other values of the parameter.

Within this framework, I consider two nested theories of belief updating: the first one is a dogmatic modeler adapted from @Heidhues2018,The second one is a switcher, as in @Ba2023. The dogmatic modeler can be seen as a switcher with an infinitely sticky initial belief so the two theories are nested. However, outside of the limiting case, they produce different 
predictions about the agent's behavior. I discuss each of them separately in what follows.

@@

Take the action space to be the same as the state space $\Omega$ and let $\Omega=A:=[\underline{a}, \infty)$.

A dogmatic agent is characterized by a learning model where he does not update their beliefs about $\theta$, instead, he holds a degenerate belief that places all probability on $\hat{\theta}$, which is potentially misspecified. 
In this case, no matter how much information he gathers against being of type $\hat{\theta}$, he will not update his beliefs about it. 
Any discrepancies between the observed outcomes and his believed type are incorporated using the Bayes rule to update his beliefs about $\omega$. @Heidhues2018 show that, under certain assumptions on the per-period utility,\footnote{The assumptions are that $u$ is twice continuously differentiable with: (i)$u_{aa}<0$ and $u_a(\underline{a},  \theta, \omega)>0>u_a(\Bar{a}, \theta, \omega)$, (ii) $u_{\theta}, u_{\omega}>0$ and (iii) $u_{a\theta}<0$ and $u_{a\omega}>0$. Where $\underline{a}$ is the
minimal action and $\Bar{a}$ is the maximal one. The direction of the derivatives is a normalization
and the results would hold even when the signs are reversed.} a dogmatic modeler will inevitably fall into a self-defeating equilibrium. The equilibrium will be such that the outcomes they observe reinforce their belief on $\omega$ in such a way that as $t\to\infty$ the agent will be sure that the state is some $\omega^{'}$ consistent with their believed type and the observed data. In other words, they will be in a self-defeating equilibrium with a stable belief that places probability one on the incorrect parameters $(\hat{\theta}, \omega_{\infty})$

The mechanism by which the dogmatic agent falls into the self-defeating equilibrium is the following: Suppose the agent holds the misspecified belief that they are type $\hat{\theta}>\theta^*$. For any prior over $\omega$, the agent will be disappointed by the outcome. That is: 
$$\bar{h}(\theta^*, \omega^*, a_t^*) := \frac{1}{T}\sum_t \left[h(\theta^*, \omega^*, a_t^*)+\varepsilon_t\right]<\frac{1}{T}\sum_t \left[h(\hat{\theta}, \mathbb{E}(\omega), a_t^*)+\varepsilon_t\right]$$
Where the right-hand-side of the inequality represents the average outcome that the agent expected.

There are two possible sources for the disappointment: the first is that the realized state is lower than the expected state; the second source is that the agent is of type $\theta^*$ and therefore, for all possible states, his gain will be lower than what he expected.

Because the agent is dogmatic, he will not update his beliefs about $\theta$ and as a consequence will attribute the disappointment to the state being lower than expected. He will continue to update in this way until he converges to a belief about $\omega$ that is stable. Such a belief will explain the observed utility perfectly and allow the agent to rationalize his dogmatic belief about $\theta$. Under the assumptions of @Heidhues2018, there is a unique value of $\omega$ at which the belief is stable, I will refer to such value as $\omega_\infty$. This mechanism is illustrated in Example 1.

@@

Where $h^t$ is the history of outcomes up to time $t$, and $\alpha$ is the switching threshold.\footnote{Notice that if $\alpha \to \infty$, the behavior of the switcher will be indistinguishable from that of the Dogmatic modeler. 
In this sense, the switcher is a generalization of the dogmatic type.} By keeping track of the 
likelihood ratio, the agent can perform a \emph{Bayesian hypothesis test} and adopt the Dogmatic belief that best fits 
the data.\footnote{In a related problem \citet{Schwarstein2021} proposes a similar updating procedure which relies on the Bayesian hypothesis test. However, in their model there is a sender who optimally chooses to propose a model that fits the data}

@@

A Bayesian agent simultaneously updates their beliefs about $\theta$ and $\omega$ by using Bayes' rule. 
The posterior at period t+1 about $\theta$ after observing outcome $o$ is given by:
$$
p_{t+1}(\theta, \omega| h^t) = \frac{p[o|\theta, \omega]p_t(\theta, \omega)}{\sum_{(\theta', \omega')}p[o|\theta', \omega']p_t(\theta', \omega')}
$$

@@

Bayesian agents choose the effort level that maximizes their expected flow payoff by taking expectations over
their prior beliefs about $\theta$ and $\omega$. Since agents are myopic, even though all the parameters 
could be identified with enough variation in choices, there will be instances in which the Bayesian will not
learn even with infinite amounts of data. This happens because, by being myopic, they do not internalize 
the trade off between flow payoff and learning. This can result in too little experimentation to learn their true type. An alternative to this approach is given by @Hestermann2021 and is discussed in the concluding remarks. Regardless, this approach is useful as a benchmark for the other theories discussed in this paper.

@@

A Self-Serving Bayesian is an agent who uses a biased version of Bayes rule to update their beliefs. They will update their beliefs about the state $\omega$ and his type $\theta$ simultaneously by over-attributing successes to a high value of $\theta$ and under-estimating the role of higher $\omega$. Similarly, he will attribute failure to a low state to a greater degree than an unbiased agent would. This sort of biased updating is referred to as \emph{misattribution bias}.

To model the misattribution bias, I take the approach of @benjamin2019, where the posterior odds are given by:

@@

with $c(\theta_H, \omega, o)<c(\theta_M, \omega, o)<c(\theta_L, \omega, o)$
and $c(\theta, \omega_L, o)<c(\theta, \omega_M, o)<c(\theta, \omega_H, o)$. And where $h_t$ is the history of signal realizations observed up to period t.

@@

Define the distortion in the likelihood by $\Delta_c p(h^t|\theta, \omega) = p(h^t|\theta, \omega)^{c(\theta, \omega, o_t)} - p(h^t|\theta, \omega)$. Notice that because $p(h^t|\theta, \omega)<1$, if $c(\cdot)<1$, a smaller exponent $c$ implies a larger distortion upwards for fixed $h^t$, and fixed parameters $\theta$ and $omega$. While a $c(\cdot)>1$ implies a distortion downwards and it is larger when $c(\cdot)$ is larger.

@@

In order to compare the predictions of the theories discussed above, I develop a unifying example where the forces behind each of the theories can be isolated. The example is a modified from @Heidhues2018 and is adapted to be implementable in the lab.

@@

Denote the true type by $\theta^*$ and the true state by $\omega^*$. The agent holds some prior belief about 
$\theta$\footnote{which is potentially misspecified as in the dogmatic and switcher cases discussed above} 
and chooses gamble $e in \{e_L, e_M, e_H\}$. The agent observes whether the gamble is a success or a failure and 
gets a payoff of $1$ and if it is a success; they get $0$ otherwise.

@@

In this section, I outline the behavior that is predicted by each of the theories discussed above. Some of the predictions are made relative to the Bayesian benchmark. In particular, the reaction to good and bad news is very informative of which model is most likely to be underlying the behavior. Define \emph{good news} to be a sample of signal realizations such that the number of successes in the sample is larger than the average number of successes observed so far. i.e. if

$$ \sum_i \mathbb{I}\{o_{iT}=success\} \geq \frac{1}{T} \sum_{t=0}^T\sum_i \mathbb{I}\{o_{it}=success\}$$
where T is the current period and $o_{t}=[o_{it}]_{i=1}^I$ is the sample observed at period $t$. \emph{Bad news} are any other kind of signal realization. Denote the set of good news at period t by $\mathcal{G}_t$ and let $s_t$ be the number of successes in $o_t$.

Given the definition of good and bad news, the reaction of the agent to each kind of news is the change in their choice, conditional on the signal realization. Therefore, what we are interested are the coefficients of the following regression model, where the actions are those that a bayesian agent would choose:

$$ \Delta a = \beta_0 + \beta_1\mathbb{I}\{o_t\in \mathcal{G}_t\} + \beta_2(s_{t} - s_{t-1}) + \beta_3\mathbb{I}\{o_t\in \mathcal{G}_t\}(s_{t} - s_{t-1})  $$

Where $(s_{t} - s_{t-1})$ is the size of the surprise and $\Delta a$ is the change in action. $\beta_2$ is the reaction of the agent to bad news with surprise $(s_{t} - s_{t-1})$. And $\beta_2 + \beta_3$ is the reaction to good news with surprise $(s_{t} - s_{t-1})$. Notice that for a dogmatic modeler $\beta_2\geq0$ and $\beta_2+\beta_3\geq 0$ as more successes are always indicative of a better state.

@@

In term of the dogmatic's reagtion to good and bad news, let $\beta^D$ denote the coefficients of the regression model above when the actions are those predicted by the dogmatic model.

**Prediction 2D:** _Before falling into a trap, the dogmatic modeler reacts more to good and bad news than the Bayesian would. i.e. $\beta_2^D>\beta_2$ and $\beta_3^D > \beta_3$._

@@

**Prediction 1S:** _With positive probability, the switcher will escape the self-defeating equilibrium and learn the true state._

Let $\beta^S$ be the coefficients of the regresion model when the actions are those taken by a switcher.

**Prediction 2S:** _When there is no change of paradigm, the switcher behaves as the dogmatic. When there is a change in paradigm, the reaction to good and bad news is attenuated or even in the opposite direction of what a Bayesian would have done. that is $\beta_2^S<\beta_2$ and $\beta_2^S + \beta_3^S<\beta_2+\beta_3$_

One caveat is that when the switcher and the dogmatic agent both start with a correctly specified belief, neither of them 
will fall into the self-defeating equilibria and thus will look identical even in the long run. This means that
in order to distinguish between the two theories, I need to look at cases where the agent starts with a misspecified belief and the switcher changes paradigm.

@@

Let $\beta^A$ be the coefficients of the regression model when the agent updates with misattribution bias.

**Prediction 1A:** _After observing a failure, the agent will adjust their effort downwards by more than a Bayesian would have done. After observing a success, the agent will adjust their effort upwards by less than a Bayesian would have done. If the bias is large enough, the agent might not adjust their effort upwards at all, or even decrease their effort in response to a success. That is $\beta_2^A > \beta_2$ and $\beta_2^A + \beta_3^A<\beta_2 + \beta_3$_

In addition to their behavior, there is an additional characteristic feature of misattribution bias that is not present in other models. This is the only model in which an agent could start off with a correct belief and become more overconfident. In the long-run.

@@

@@

$$ arg\min_j \sum_k \frac{1}{T}\sum_t{\left[e_{ikt}-\tilde{e}_{t}(\theta, \omega, j)\right]^2} $$
The histogram in Figure \ref{fig:best-models} shows the distribution of the best fitting models in the sample. The average fit of each model is reported in the Appendix.\footnote{The best fitting models are robust to different distance functions. In particular, the share of subjects best explained by misattribution bias is the same when using norm 1.}

@@

A different bias that has been studied and suggested in the literature is the confirmation bias. In particular,
it has been suggested that subjects who are overestimate $\theta$ in their prior, will over-attribute 
successes to a high type and under-attribute them to a low type. Symmetrically, subjects who initially 
underestimate $\theta$ will over-attribute failures to a low type and under-attribute them to a high type.
I do not find evidence of such a bias in the data. This is consistent with what @Mobius2022 find in their
experiment.

I further investigate possible heterogeneity in the underlying models being used by estimating the 
structural parameters of the models. I find that under the average parameters, most subjects are best explained by misattribution bias in both treatments. However, for the subjects for whom other models are a better explanation, those models have a better fit than the misattribution bias does for those better explained by it.
